NDNS multilevel LCA CW2CB2 or CW2CB3

### 2018-08-15 paper/Chapters/Chapter4.tex
eating patterns discusses

Among the dietary diaries collected, there were three distinct latent classes specifically for carbohydrate intake: 1) high probabilities of having high carbohydrate contained food across the hours of day (high percentage carbohydrate day); 2) low carbohydrate food dominant through out the hours of day (low percentage carbohydrate day); and 3) always having lunch and dinner day (regular meals day). And from these day level classifications and their features, one might anticipate that individuals who managed to follow the regular meals day might be eating a relatively healthier diet because it seemed to be a regular temporal eating habit; at this time point, we also believed that those who followed either high or low carbohydrate percentage days would probably consume higher total energy than those who followed most regular meals days.

However, when the MLCA extended the model to individual level, three types of persons were further defined depending on their 4-day-diary: 1) low carbohydrate eaters, who mostly followed "regular meals day"; 2) moderate carbohydrate eaters, who had similar probabilities of following either "regular meals day" or "high percentage carbohydrate day"; 3) high carbohydrate eaters, who followed "high percentage carbohydrate day" for half of their survey. For the first time, as far as we know, the day-to-day food intake pattern variation within individuals was successfully captured by MLCA models. Results from the MLCA models showed that from the perspective of carbohydrate consumption, people were indeed changing their diet from day to day even within a short term period of survey. The MLCA models allowed the probability of following a certain type of carbohydrate eating day to vary across individuals. This properly accounted for the fact that for some people, their probability of following a type of food eating pattern during the survey could be higher/lower than that in the others. This finding also suggested that assuming a person will following a certain type of food intake pattern is not appropriate.

Surprisingly, low carbohydrate eaters whose dietary recordings suggested that they were mostly following a regular temporal meals pattern turned out to consume the highest amount of total energy among the three type of carbohydrate eaters. Detailed profiling of energy composition according to the time slots revealed that low carbohydrate eaters actually had higher proportion of energy contributed by both alcohol and fat. High percentage of fat consumption was shown in all 7 time slots, energy coming from alcohol exceeded more than one fifth of the total energy after 8 pm. These findings explained why they were actually consuming the highest energy among the three types of carbohydrate eaters. However, we also found that participants consuming low carbohydrate had higher prevalence of diabetes, hypertension, and obesity. These health issues might possibly lead them (or advices were given from their physicians) to replace carbohydrates in their diet to other energy sources such as fat, protein, or even alcohol. Therefore, there is a possibility that they chose to follow low carbohydrate diets out of healthy purpose, but many of them were replacing carbohydrates with higher energy condensed food or even alcohol at night which might indeed be a public health concern.

Further more, looking into the details of the timing and composition of the energy intake among the moderate carbohydrate eaters. We realised that although moderate carbohydrate eaters did not consume as much alcohol as low carbohydrate eaters at night. But they were

>>>>

Among the dietary diaries collected, there were three distinct latent classes specifically for carbohydrate intake: 1) high probabilities of having high carbohydrate contained food across the hours of day (high percentage carbohydrate day); 2) low carbohydrate food dominant through out the hours of day (low percentage carbohydrate day); and 3) always having lunch and dinner day (regular meals day). And from these day level classifications and their features, one might anticipate that individuals who managed to follow the regular meals day might be eating a relatively healthier diet because it seemed to be a regular temporal eating habit; at this time point, we also believed that those who followed either high or low carbohydrate percentage days would probably consume higher total energy than those who followed mostly regular meals days.

However, when the MLCA extended the model to individual level, three types of persons were further defined depending on their 4-day-diary: 1) low carbohydrate eaters, who mostly followed "regular meals day"; 2) moderate carbohydrate eaters, who had similar probabilities of following either "regular meals day" or "high percentage carbohydrate day"; 3) high carbohydrate eaters, who followed "high percentage carbohydrate day" for half of their survey. For the first time, as far as we know, the day-to-day food intake pattern variation within individuals was successfully captured by MLCA models. Results from the MLCA models showed that from the perspective of carbohydrate consumption, people were indeed changing their diet from day to day even within a short term period of survey. The MLCA models allowed the probability of following a certain type of carbohydrate eating day to vary across individuals. This properly accounted for the fact that for some people, their probability of following a type of food eating pattern during the survey could be higher/lower than that in the others. This finding also suggested that assuming a person will always follow a certain type of food intake pattern is not appropriate.

Surprisingly, low carbohydrate eaters whose dietary recordings suggested that they were mostly following a regular temporal meals pattern turned out to consume the highest amount of total energy among the three types of carbohydrate eaters. Detailed profiling of energy composition according to the time slots revealed that low carbohydrate eaters actually had higher proportion of energy contributed by both alcohol and fat. High percentage of fat consumption was shown in all 7 time slots, energy coming from alcohol exceeded more than one fifth of the total energy after 8 pm. These findings explained why they were actually consuming the highest energy among the three types of carbohydrate eaters. However, we also found that participants consuming low carbohydrate food had higher prevalence of diabetes, hypertension, and obesity. These health issues might possibly lead them (or advices were given from their physicians) to replace carbohydrates in their diet to other energy sources such as fat, protein, or even alcohol. Therefore, there is a possibility that they chose to follow low carbohydrate diets out of health purpose, but many of them were replacing carbohydrates with higher energy condensed food or even alcohol at night which might indeed be a public health concern.

Next, when looking into the details of the timing and composition of the energy intake among the moderate carbohydrate eaters, we realised that although these individuals did not consume as much alcohol as low carbohydrate eaters at night, they consumed the highest amount of energy, especially during time period as late as after 10 pm. People fell into moderate carbohydrate eaters group seemed to have the tendency of having their food or meals later than the other two types of carbohydrate eaters. They consumed the highest amount of carbohydrates and also total energy during the following time slots: 9-12 noon, 2-5 pm, 8-10 pm, and 10pm-6am. These individuals in the NDNS RP were younger, mostly single, with lower average income, and lower education level. They might possibly correspond to the "late eaters" defined by previous studies \parencite{leech2017temporal, Mansukhani2018}.

Lastly, the high carbohydrate eaters identified by our MLCA models had the highest absolute total amount of carbohydrate intake. Most of their energy intake occured during 6-9 am, 12-2 pm, and 5-8 pm time slots while their average energy consumption was not the highest (between low and moderate carbohydrate eaters). High carbohydrate eaters were also found to be the ones that consumed the least amount of energy after 8 pm. Therefore, contrary to what was anticipated at the beginning, people who followed high percentage carbohydrate days for most of their time, were actually eating a healthier diet compared with the other two eating patterns.

Our analyses looking for different temporal carbohydrate also highlighted the complexity of eating pattern behaviours in the population and the utility of exploratory, data-driven methods to objectively identify eating patterns that reflect both timing and quantities of food intake, which may have not been detected so far in the literature.

### 2018-08-15 paper/main.tex
eating patterns discusses

\thesistitle{The time patterns of carbohydrate intake in the UK adults -- the National Dietary and Nutrition Survey 2008/09-15/16 programme} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
\supervisor{Professor Luigi \textsc{Palla}} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
\examiner{} % Your examiner's name, this is not currently used anywhere in the template, print it elsewhere with \examname
\degree{MSc in Medical Statistics} % Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
\author{Chaochen \textsc{Wang}} % Your name, this is used in the title page and abstract, print it elsewhere with \authorname
\addresses{} % Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname

@@

I would like to thank the team of the National Dietary and Nutrition Survey Rolling Programme (NDNS RP) who have made their data available to the public for academic study. \\

@@

\begin{abstract}
\addchaptertocentry{\abstractname} % Add the abstract to the table of contents
The National Dietary and Nutrition Survey (NDNS) database of detailed four-day food diaries was used to \dots
\end{abstract}

@@

\textbf{AIC} & \textbf{A}kaike \textbf{I}information \textbf{C}riterion \\
\textbf{A1c} & Haemoglobin \textbf{A1c}: Glycated hemoglobin \\
%\textbf{aBIC} & \textbf{a}djusted \textbf{B}ayesian \textbf{I}nformation \textbf{C}riterion \\
\textbf{BMI} & \textbf{B}ody \textbf{M}ass \textbf{I}ndex \\
%\textbf{cAIC} & \textbf{c}onsistent \textbf{A}kaike \textbf{I}nformation \textbf{C}riterion \\
\textbf{BIC} & \textbf{B}ayesian \textbf{I}information \textbf{C}riterion \\
\textbf{CI}  & \textbf{C}onfidence \textbf{I}nterval \\
\textbf{DM}  & \textbf{D}iabetes \textbf{M}ellitus \\
\textbf{EM}  & \textbf{E}xpectation \textbf{M}aximazation  \\
\textbf{FSA} & \textbf{F}ood \textbf{S}tandards \textbf{A}gency  \\
\textbf{HDL} & \textbf{H}igh \textbf{D}ensity \textbf{L}ipoprotein cholesterol \\
\textbf{LDL} & \textbf{L}ow \textbf{D}ensity \textbf{L}ipoprotein cholesterol \\
\textbf{LCA} & \textbf{L}atent \textbf{C}lass \textbf{A}nalysis\\
%\textbf{LCGA} & \textbf{L}atent \textbf{C}lass \textbf{G}rowth \textbf{A}nalysis \\
%\textbf{LTA} & \textbf{L}atent \textbf{T}ransition \textbf{A}nalysis \\
\textbf{MAFF} & \textbf{M}inistry of \textbf{A}griculture, \textbf{F}isheries and \textbf{F}ood \\
%\textbf{MAR} & \textbf{M}issing \textbf{A}t \textbf{R}andom \\
%\textbf{MCAR} & \textbf{M}issing \textbf{C}ompletely \textbf{A}t \textbf{R}andom \\
\textbf{MLCA} & \textbf{M}ultilevel \textbf{L}atent \textbf{C}lass \textbf{A}nalysis \\
%\textbf{MNAR} & \textbf{M}issing \textbf{N}ot \textbf{A}t \textbf{R}andom \\
\textbf{ML} & \textbf{M}aximum \textbf{L}ikelihood \\
\textbf{NDNS} & the \textbf{N}ational \textbf{D}ietary and \textbf{N}utrition \textbf{S}urvey \\
\textbf{OR} & \textbf{O}dds \textbf{R}atio \\
\textbf{PHE} & \textbf{P}ublic \textbf{H}ealth \textbf{E}ngland  \\
\textbf{PSUs} & \textbf{P}rimary \textbf{S}ampling \textbf{U}nits \\
\textbf{TC} & \textbf{T}otal \textbf{C}holesterol \\
\textbf{TG} & \textbf{T}ri\textbf{g}lycerides \\
\textbf{WC}   & \textbf{W}aist \textbf{C}ircumference \\
\end{abbreviations}

>>>>

\thesistitle{The time patterns of carbohydrate intake in the UK adults -- the National Diet and Nutrition Survey 2008/09-15/16 programme} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
\supervisor{Professor Luigi \textsc{Palla}} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
\examiner{} % Your examiner's name, this is not currently used anywhere in the template, print it elsewhere with \examname
\degree{MSc in Medical Statistics} % Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
\author{Chaochen \textsc{Wang}} % Your name, this is used in the title page and abstract, print it elsewhere with \authorname
\addresses{} % Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname

@@

I would like to thank the team of the National Diet and Nutrition Survey Rolling Programme (NDNS RP) who have made their data available to the public for academic study. \\

@@

\begin{abstract}
\addchaptertocentry{\abstractname} % Add the abstract to the table of contents
The National Diet and Nutrition Survey (NDNS) database of detailed four-day food diaries was used to \dots
\end{abstract}

@@

\textbf{AIC} & \textbf{A}kaike \textbf{I}information \textbf{C}riterion \\
\textbf{A1c} & Haemoglobin \textbf{A1c}: Glycated hemoglobin \\
%\textbf{aBIC} & \textbf{a}djusted \textbf{B}ayesian \textbf{I}nformation \textbf{C}riterion \\
\textbf{BMI} & \textbf{B}ody \textbf{M}ass \textbf{I}ndex \\
%\textbf{cAIC} & \textbf{c}onsistent \textbf{A}kaike \textbf{I}nformation \textbf{C}riterion \\
\textbf{BIC} & \textbf{B}ayesian \textbf{I}information \textbf{C}riterion \\
\textbf{CI}  & \textbf{C}onfidence \textbf{I}nterval \\
\textbf{DM}  & \textbf{D}iabetes \textbf{M}ellitus \\
\textbf{EM}  & \textbf{E}xpectation \textbf{M}aximazation  \\
\textbf{FSA} & \textbf{F}ood \textbf{S}tandards \textbf{A}gency  \\
\textbf{HDL} & \textbf{H}igh \textbf{D}ensity \textbf{L}ipoprotein cholesterol \\
\textbf{LDL} & \textbf{L}ow \textbf{D}ensity \textbf{L}ipoprotein cholesterol \\
\textbf{LCA} & \textbf{L}atent \textbf{C}lass \textbf{A}nalysis\\
%\textbf{LCGA} & \textbf{L}atent \textbf{C}lass \textbf{G}rowth \textbf{A}nalysis \\
%\textbf{LTA} & \textbf{L}atent \textbf{T}ransition \textbf{A}nalysis \\
\textbf{MAFF} & \textbf{M}inistry of \textbf{A}griculture, \textbf{F}isheries and \textbf{F}ood \\
%\textbf{MAR} & \textbf{M}issing \textbf{A}t \textbf{R}andom \\
%\textbf{MCAR} & \textbf{M}issing \textbf{C}ompletely \textbf{A}t \textbf{R}andom \\
\textbf{MLCA} & \textbf{M}ultilevel \textbf{L}atent \textbf{C}lass \textbf{A}nalysis \\
%\textbf{MNAR} & \textbf{M}issing \textbf{N}ot \textbf{A}t \textbf{R}andom \\
\textbf{ML} & \textbf{M}aximum \textbf{L}ikelihood \\
\textbf{NDNS} & the \textbf{N}ational \textbf{D}iet and \textbf{N}utrition \textbf{S}urvey \\
\textbf{OR} & \textbf{O}dds \textbf{R}atio \\
\textbf{PHE} & \textbf{P}ublic \textbf{H}ealth \textbf{E}ngland  \\
\textbf{PSUs} & \textbf{P}rimary \textbf{S}ampling \textbf{U}nits \\
\textbf{TC} & \textbf{T}otal \textbf{C}holesterol \\
\textbf{TG} & \textbf{T}ri\textbf{g}lycerides \\
\textbf{WC}   & \textbf{W}aist \textbf{C}ircumference \\
\end{abbreviations}
